\Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project

### 2014-06-12 articles/2014/whisper/aricleEbauche/articleIrodsUserMeeting.tex
small change syntax spelling

The \emph{Whisper} \footnote{FP7 ERC Advanced grant 227507, see whisper.obs.ujf-grenoble.fr} 
project is a an european project on seismology whose goal is to study properties of the earth
with the seismic ambient noise such that evolution of seismic waves speed. 
This noise is almost all the signal continuously recorded by
the seismic stations worldwide (Europe, China, USA, Japan), except earthquakes. It offers new
observables for the seismologists, new types of virtuals seismograms that are not only located at the 
place of earthquakes. For instance, one can obtain wave paths that probes the deepest part of the Earth.
%(variations in space and time of the seismic waves)

@@

In addition, the calculation operations downstream may produce even more data than the observation data.
To give an order of magnitude, more than 200 TB has been managed by the Whisper project at the same time. 
A classical processing produces 8 TB in 5 days. An other one 'read' 3 or 4 TB and 'produced' 1 TB in 6 hours.
Many tests of the signal processing are done and computational data are deleted as and when required.
\\

@@

A part of the Whisper project is specificaly dedicated to the IT codes. This includes to design a specification, an implementation and some optimisations of
a sequence of a data-management and of a computations \footnote{see code-whisper.isterre.fr/html/ (part of the design)}.
This project uses some own IT resources (servers, dedicated bay) but also uses common IT infrastructure of the university.
We developed also adaptations for the IT infrastrutures and we provide technical support for researchers.

Most of the IT codes are writing with the \emph{Python} language and uses intensively the scientific libraries \emph{Scipy} (fortran and C embeded) 
and \emph{Obspy} \footnote{ see www.obspy.org} (essentially the 'read' function). The latest one is dedicated to the seismological community.

The IT codes consists of several tools described schematically at the figure \ref{whisperGeneralWorflow} and grouped into three parts. The first one concerns the signal processing, the second part permits the computation of the correlations and the last part consists of codes for the analysis of the correlations.

A first package provides a flexible way to process raw data, to specify a pipeline of pre-processing of signal.
The user starts by specifying a directory, a set of seismic stations and a set of dates. Then the codes scan the directory
and extracts all pieces of seismograms (also called traces) and rearranges them in a specific architecture of files in order to calculate the correlations to the next step.
We use here intensively the function 'read' of the library Obspy which allows to open most of format files seismogram.
The user also define his own sequence of processings. He can use the functions predefined but also the Python libraries he needs. 
Moreover he can add eventually his own processing.

The second package concerns correlations. 
Roughly speaking, a correlation is an operation with two seismograms (for a given time windows) that provides the coherent part of the two seismograms 
(associated to the 2 stations) which is the seismic waves that propagate between the 2 stations.
(Moreover, in some favorables cases, it converges to the Green's function). 
Thus, the code computes all the correlations and provides an architecture of files that corresponds to all the couples of seismograms (for each date).

@@

Nevertheless, the main optimization was made by testing the behaviour of the carbage collector of Python in order to follow the cache heuristics. More precisely, we do not use the gc module or the 'del' statement but we try to schedule and localize the line of code in order to find the good unfolding that uses the architecture optimally.

The last part of computer codes concerns the analysis of correlations (the virtual new seismograms) with methods such as beamforming, doublet or inversion.
We also compute correlations of correlations C3 (also new seismograms). For example, we study the variations
in velocity of seismic waves as we illustrate below.

@@

The data-intensive processing
needs obviously an IT infrastructure in order to couple storage and computation.
In our cases, most of the processing are embarrassingly parallel.
The amount of data and the location of compute nodes available suggests using a 
distributed storage system and a grid manager.

>>>>

The \emph{Whisper} \footnote{FP7 ERC Advanced grant 227507, see whisper.obs.ujf-grenoble.fr} 
project is a an european project on seismology whose goal is to study properties of the earth
with the seismic ambient noise such that evolution of seismic waves speed. 
This noise is almost all the signal continuously recorded by
the seismic stations worldwide (Europe, China, USA, Japan), except earthquakes. It offers new
observables for the seismologists, new types of virtuals seismograms that are not only located at the 
place of earthquakes and that are provided by the operation of correlation  which requires significant computations.
For instance, one can obtain wave paths that probes the deepest part of the Earth.
%(variations in space and time of the seismic waves)

@@

In addition, the calculation operations downstream may produce even more data than the observation data.
To give an order of magnitude, more than 200 TB have to be processed by the Whisper project at the same time. 
A classical processing produces 8 TB in 5 days. Another computation 'read' 3 or 4 TB and 'produced' 1 TB in 6 hours.
Many tests of the signal processing are done and computational data are deleted as and when required.
\\

@@

A part of the Whisper project is specificaly dedicated to the IT codes. This includes to design a specification, an implementation and some optimisations of the sequence of data-management as well as of the computations \footnote{see code-whisper.isterre.fr/html/ (part of the design)}.
This project uses its own IT resources (servers, dedicated bay) as well as the common IT infrastructure of the university.
We developed also adaptations for the IT infrastrutures and we provide technical support for researchers.

Most of the IT codes are written with the \emph{Python} language and make intensive use of the scientific libraries \emph{Scipy} (fortran and C embeded) 
and \emph{Obspy} \footnote{ see www.obspy.org} (essentially the 'read' function) which is dedicated to the seismological community.

The IT codes consist of several tools described schematically in figure \ref{whisperGeneralWorflow} and grouped into three parts. The first one concerns the signal processing, the second part permits the computation of the correlations and the last part consists of codes for the analysis of the correlations.

The first package provides a flexible way to process raw data, to specify a pipeline of pre-processing of signal.
The user starts by specifying a directory, a set of seismic stations and a set of dates. Then the codes scans the directory
and extracts all pieces of seismograms (also called traces) and rearranges them in a specific architecture of files in order to calculate the correlations to the next step.
We use here intensively the function 'read' of the library Obspy which allows to open most of the seismogram file formats.
The user also define his own sequence of processings. He can use the functions predefined but also the Python libraries he needs and,
moreover, he can add eventually his own functions.

The second package concerns correlations. 
Roughly speaking, a correlation is an operation with two seismograms (for a given time window) that provides the coherent part of the two seismograms 
(associated to the 2 stations) which is the seismic waves that propagate between the 2 stations.
(Moreover, in some favorables cases, it converges to the Green's function). 
Thus, the code computes all the correlations and provides an architecture of files that corresponds to all the couples of seismograms (for each date).

@@

Nevertheless, the main optimization was made by testing the behaviour of the carbage collector of Python in order to follow the cache heuristics. More precisely, we do not use the 'gc' module or the 'del' statement but we try to schedule and localize the line of code in order to find the good unfolding that uses the architecture optimally.

The last part of computer codes concerns the analysis of correlations (the virtual new seismograms) with methods such as beamforming, doublet or inversion.
We also compute correlations of correlations C3 (also new seismograms). For example, we study the variations
in velocity of seismic waves as we illustrate below in figure \ref{JapanVelocity}.

@@

The data-intensive processing
needs obviously an IT infrastructure in order to couple storage and computation.
In our cases, most of the processings are embarrassingly parallel.
The amount of data and the location of available compute nodes suggests the use of a 
distributed storage system and a grid manager.
